### 2018-07-31 CS761-RandomizedAlgorithms/assignments/a4.tex
q2 of a4

Consider an instance $I$ of $k$SAT with $n$ variables $x_1,\dots,x_n$ and $m$ clauses $C_1, \dots, C_m$. We solve it using the following generalization of Sch\"oning's Algorithm:

@@

		\item Repeat up to $3n$ steps, terminating if all clauses satisfied: switch the value of a random variable in an unsatisfied clause.

@@

Suppose we have assignment $A_i$ at iteration $i$ with $j$ mismatches with $S$, i.e. $X_i = n-j$. Then we reach state $n$ in $3j$ steps provided there are $j$ steps back and $2j$ steps forward among these $3j$ steps. This happens with probability
$${3j\choose j}(\frac{k-1}{k})^j(\frac{1}{k})^{2j}.$$
This is a lower bound on reaching $n$ in $3n$ steps.

@@

$$\sqrt{2\pi m}\frac{m}{e}^m \leq m! \leq 2\sqrt{2\pi m}\frac{m}{e}^m$$

@@

$${3j\choose j} \geq \frac{\sqrt{2\pi 3j}}{4\sqrt{2\pi 2j}\sqrt{2\pi j}}\frac{3j}{e}^{3j}\frac{e}{2j}^{2j}\frac{e}{j}^j = \frac{c}{\sqrt{j}}\frac{27}{4}^j$$
where $c = \frac{\sqrt{3}}{8\sqrt{\pi}}$ (a constant).

@@

$$\frac{c}{\sqrt{j}}\frac{27}{4}^j(\frac{k-1}{k})^j(\frac{1}{k})^{2j}$$

>>>>

Consider an instance $I$ of $k$SAT with $n$ variables $x_1,\dots,x_n$ and $m$ clauses $C_1, \dots, C_m$. We may assume that $k$ is odd, otherwise replace $k$ with $k+1$ and add a dummy variable to each clause of $I$ if necessary. Solve $I$ using the following generalization of Sch\"oning's Algorithm:

@@

		\item Repeat up to $kn$ steps, terminating if all clauses satisfied: switch the value of a random variable in an unsatisfied clause.

@@

Suppose we have assignment $A_i$ at iteration $i$ with $j$ mismatches with $S$, i.e. $X_i = n-j$. Let $a = (k-1)/2$. We reach state $n$ in $kj$ steps provided there are $aj$ steps back and $(a+1)j$ steps forward among these $kj$ steps. This happens with probability
$${kj\choose j}(\frac{k-1}{k})^{aj}(\frac{1}{k})^{(a+1)j}.$$
This is a lower bound on reaching $n$ in $kn$ steps.

@@

$$\sqrt{2\pi m}(\frac{m}{e})^m \leq m! \leq 2\sqrt{2\pi m}(\frac{m}{e})^m$$

@@

$${kj\choose j} \geq \frac{\sqrt{2\pi kj}}{4\sqrt{2\pi aj}\sqrt{2\pi (a+1)j}}(\frac{kj}{e})^{kj}(\frac{e}{aj})^{aj}(\frac{e}{(a+1)j})^{(a+1)j} = \frac{c}{\sqrt{j}}(\frac{k^k}{a^a(a+1)^{a+1}})^j$$
where $c = \frac{\sqrt{k}}{4\sqrt{2\pi a(a+1)}}$ (a constant when $k$ is constant).

@@

$$\frac{c}{\sqrt{j}}(\frac{k^k}{a^a(a+1)^{a+1}})^j(\frac{k-1}{k})^{aj}(\frac{1}{k})^{(a+1)j} = \frac{c}{\sqrt{j}}\frac{(k-1)^{aj}}{a^{aj}(a+1)^{(a+1)j}} = \frac{c}{\sqrt{j}}\frac{(2a)^{aj}}{a^{aj}(a+1)^{(a+1)j}},$$
and hence we obtain a lower bound on the probability of reaching $n$ from $j$ mismatches of
$$\frac{c}{\sqrt{j}}\frac{2^{aj}}{(a+1)^{(a+1)j}}.$$
Therefore the probability of the algorithm finding a satisfying assignment in $kn$ steps is at least
\sum_{j=0}^n\frac{c}{\sqrt{j}}\frac{2^{aj}}{(a+1)^{(a+1)j}} {n\choose j}\frac{1}{2^n} &\geq \frac{c}{\sqrt{n}}\frac{1}{2^n}\sum_{j=0}^n{n\choose j}\frac{2^{aj}}{(a+1)^{(a+1)j}} \\
&=\frac{c}{\sqrt{n}}\frac{1}{2^n}(1+\frac{2^a}{(a+1)^{a+1}})^n &\text{by the binomial theorem}
Hence the expected runtime of the algorithm is at most
$$\frac{\sqrt{n}2^n}{c}(1+\frac{2^a}{(a+1)^{a+1}})^{-n},$$
a function of $n$ and $k$ (since $a$ and $c$ are functions of $k$). When $k\geq 3$ is constant, $\frac{2^a}{(a+1)^{a+1}}$ is a positive constant smaller than $1$, call it $d$. Then our expected runtime is at most
$$O(\sqrt{n}(\frac{2}{d})^n) = o(\sqrt{n}2^n).$$
